Measurement of a dwarf galaxy rotational velocity from its Lyman-alpha line emission

### 2016-08-03 paper/letter.tex
abstract

\begin{document}
\pagestyle{empty}
\noindent
\textbf{Þú ert jörðin}
\\
\\
JEFR$^{1}$, MCRG$^1$, JNGC$^2$, MG$^3$, MD$^3$
\\
\\
\scriptsize
{$^1$ Bogota
\\
$^2$ Tucson
\\
$^3$ Oslo
\normalsize
\\
\\
\textbf{
  Star-forming Compact Dwarf Galaxies (CDGs) resemble the expected
  pristine conditions of the first galaxies in the Universe.    
Before the observational detection of the first galaxies becomes
reality, CDGs are the best systems to test our ideas on primordial
galaxy formation and evolution.    
Here we report on one of such CDGs, \tol, which presents
a broad symmetric Lyman-$\alpha$ emission that had evaded theoretical
interpretation so far. 
We explain these features by two different models: an
homogeneous sphere undergoing gas bulk rotation and an interstellar
medium composed by clumps with random motions.
It is the first time that an observed \lya\ spectrum can be explained
with these physical conditions.
We  find that both models independently require high velocities that
translate into a dynamical mass at least $10$ times larger than the
neutral mass hydrogen in the galaxy. 
We argue that a possibility to explain the high
dynamical mass is the presence of a super-massive black hole.
The implications of our findings for the study of LAEs, including the
first generation of galaxies are manifold. 
It demonstrates the importance of considering rotation and multiphase
physics under the possible conditions shaping the \lya\ spectra of the
first galaxies. 
Additionally, if future kinematic maps of \tol\ confirm the high
velocities postulated in our model, it would provide new
evidence for dwarf galaxies as hosts of supermassive black holes. 
}

>>>>

\begin{document}
\pagestyle{empty}
\noindent
\textbf{Þú ert jörðin}
\\
\\
JEFR$^{1}$, MCRG$^1$, JNGC$^2$, MG$^3$, MD$^3$
\\
\\
\scriptsize
{$^1$ Bogota
\\
$^2$ Tucson
\\
$^3$ Oslo
\normalsize
\\
\\
\textbf{
  Star-forming Compact Dwarf Galaxies (CDGs) resemble the expected
  pristine conditions of the first galaxies in the Universe.    
Before the observational detection of the first galaxies becomes
reality, CDGs are the best systems to test our ideas on primordial
galaxy formation and evolution.    
Here we report on one of such CDGs, \tol, which presents
a broad symmetric Lyman-$\alpha$ emission that had evaded theoretical
interpretation so far. 
We explain these features by two different models: an
homogeneous sphere undergoing gas bulk rotation and an interstellar
medium composed by clumps with random motions.
It is the first time that an observed \lya\ spectrum can be explained
assuming these physical conditions.
We  find that both models independently require high velocities that
translate into a dynamical mass at least $10$ times larger than the
neutral mass hydrogen in the galaxy. 
We argue that a possibility to explain the high
dynamical mass is the presence of a super-massive black hole. 
Ionization emission lines present in this galaxy support the idea that
\tol\ might have harbored an an Active Galactic Nucleus  
The implications of our findings for the study of LAEs, including the
first generation of galaxies are manifold. 
It demonstrates the importance of considering rotation and multiphase
physics under the possible conditions shaping the \lya\ spectra of the
first galaxies. 
Additionally, if future kinematic maps of \tol\ confirm the high
velocities postulated in our model, it would provide new
evidence for dwarf galaxies as hosts of active supermassive black
holes.  
}
